Field training 02 - Leaf traits measurements

### 2023-06-13 submission_2/submission_2.qmd
Intro, Task 3,5 and tables

# Task 1: Leaf sampling

## Procedure

Date: 16.05.2023

@@

Participants: Denis, Fabian, Louis, Lukas, Marie-Louise, Maxwell

@@

>>>>

# Leaf Properties Sampling

# Task 1: Procedure

<!-- Task1: Leaf sampling in the park and quick run through of all processing steps -->

Date: 23.05.2023

@@

Participants: Dennis, Fabian, Louis, Lukas, Marie-Louise, Maxwell

@@

Material used for Sampling:

- Paper Bags and Pencil
- Smartphone (GPS + plant species determination)
- Book (Flora Vegetativa)
- Survey sheet

Materials used for Analysis:

- Soil Plant Analysis Development (SPAD) meter for chlorophyll measurements
- Precision balance
- Leaf scanner
- A4 paper and camera
- Drying cabinet

The aim of the leaf property sampling was to collect and analyse leaf traits of different plant species. On May 23rd 2023, we collected leaf samples of eight different species in Kanonenteichanlage, a small park between Liebigstraße and Brüderstraße in 04103 Leipzig, Germany.

The plant species that were sampled were choosen based on their abundance in the park. For each species 10 leaves from different individuals were sampled. When sampling leaves we were careful to sample the whole leaf. Leaves were chosen based on the following criteria, healthy, fully grown and from a sun-lit environment where possible. We sampled trees, shrubs and herbs. The plants were identified using the smartphone app 'FloraIncognita' and the latin names of the species were recorded in a table. The leaves were collected in labeled paper bags and taken back to the seminar room for further analyisis. The leaf traits recorded for each species were fresh weight, dry weight, leaf area, and chlorophyll content. The fresh weight was determined first using a precision balance. In the next step chlorophyll concentration was derived from plant greenness using a Soil Plant Analysis Development (SPAD) meter. We used the leaf scanner to measure leaf area, we also took pictures of the leafs on an A4 sheet to derive leaf area using a script for image analysis. The leaves were then put in a drying cabinate at 50° to 55° for 48h. In a last step dry weight was determined. All the data was recorded in a google sheet.

## Leaf water content

<!-- Task 3 and 5 -->
The total fresh weight and dry weight of the leave samples were used to determine the equivalent water thickness (EWT, @tbl-data-summary). The total weight of the leaves was determined through carefully placing the samples on precision laboratory balances. The equivalent water thickness is calculated by substracting the weight of the dry sample from the weight of the fresh sample and then deviding by the product of the density of water and the leaf area (@eq-EWT).

 $$
 EWT = \frac{fresh\: weight\: [mg]-dry\: weight\: [mg]}{1000\:\frac{mg}{cm^{3}}\; \; leaf\: area\: [cm^{2}]}
 $$ {#eq-EWT}

<!-- change to landscape -->
\KOMAoptions{paper=landscape,pagesize} 
\recalctypearea

<!-- please feel free to make these tables look prettier-->

| Species | Leaf area (scanner) cm² | Leaf area (camera) cm² | Fresh weight [mg] | Dry weight [mg] | LMA [mg/cm²] scanner | LMA [mg/cm²] camera | EWT [mg/cm²] scanner | EWT [mg/cm²] camera |
|---|---|---|---|---|---|---|---|---|
| Philadelphus Coronarius | 719.97 | 751.19 | 13830 | 2440 | 19.20913371 | 18.41078822 | 0.02 | 0.02 |
| Spirae Chamaedryfolia | 128.27 | 122.26 | 1760 | 620 | 13.72105715 | 14.39555047 | 0.01 | 0.01 |
| Geranium molle | 119.11 | 141.13 | 3590 | 590 | 30.14020653 | 25.43753986 | 0.03 | 0.02 |
| Alliaria petiolata | 810.5 | 892.12 | 17720 | 3630 | 21.8630475 | 19.86279873 | 0.02 | 0.02 |
| Geum urbanum | 184.8 | 194.62 | 2390 | 860 | 12.93290043 | 12.28034118 | 0.01 | 0.01 |
| Ballota nigra | 105.88 | 110.4 | 2090 | 610 | 19.73932754 | 18.93115942 | 0.01 | 0.01 |
| Geranium sanguineum | 82.62 | 97.56 | 1910 | 530 | 23.11788913 | 19.57769578 | 0.02 | 0.01 |
| Acer platanoides | 1142.2 | 1343.79 | 15110 | 4530 | 13.22885659 | 11.24431645 | 0.01 | 0.01 |

: Leaf properties: leaf area, leaf fresh and dry weight, LMA (leaf mass per area) and EWT (equivalent water thickness) {#tbl-data-summary}

| Species | SPAD mean | SPAD stdv | chl Eq. 1 | chl Eq. 2 | chl Eq. 3 | chl Eq. 4 | chl Eq. 5 |
|---|---|---|---|---|---|---|---|
| Philadelphus Coronarius | 47.65 | 1.681434573 | 265.467675 | 621.573065 | 608.2915624 | 61.56568402 | 62.75488428 |
| Spirae Chamaedryfolia | 45.6 | 3.215932559 | 243.8084 | 584.63104 | 564.7495158 | 56.03685716 | 56.72252579 |
| Geranium molle | 36.1 | 3.919183588 | 155.5059 | 425.94494 | 386.0649413 | 36.23261347 | 35.51025089 |
| Alliaria petiolata | 27.64 | 1.416725489 | 93.583776 | 301.9521344 | 257.008503 | 24.57293509 | 23.40017292 |
| Geum urbanum | 28.03 | 2.746330886 | 96.092139 | 307.3093226 | 262.3688604 | 25.01677595 | 23.85444116 |
| Ballota nigra | 38.44 | 4.576558629 | 175.413216 | 463.1218304 | 426.6499938 | 40.34089174 | 39.8524312 |
| Geranium sanguineum | 48.9 | 6.534332407 | 279.1283 | 644.56894 | 635.7441646 | 65.20131502 | 66.7438002 |
| Acer platanoides | 31.59 | 5.872998477 | 120.535811 | 357.8139034 | 313.8933755 | 29.4575616 | 28.43095204 |

: Leaf properties: SPAD measurements and chlorophyll concentration from 5 different equations {#tbl-data-summary}

<!-- change back to portrait -->
\KOMAoptions{paper=portrait,pagesize} 
\recalctypearea
